Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2020-08-21 1.intro/introduction.tex

A common theme in such applications is that the errors are correlated, and that the signal vectors are believed to be sparse: the number of non-zero components in $\mu$ is small compared to the number of test performed --- in the cybersecurity example, while monitoring is performed over a large number of servers, it is often believed that only very few servers will be experiencing problems at any time.
Under such sparsity assumptions, it is natural to ask if and when one can reliably {(1)} detect the signals, and {(2)} recover the support set $S$.
% Further, one would like to characterize the statistical procedures that achieve efficient detection and support recovery, whenever such goals become possible to achieve.
We explore both the \emph{detection} and the \emph{support recovery} problems in this thesis. 
More precisely, we are interested in the theoretical feasibility of both problems, and seek minimal conditions under which these problems can be consistently solved in large dimensions.

@@

Both the Gaussianity assumption and the independence assumption are relaxed in Chapter \ref{chap:exact-support-recovery}.
Established are the necessary and sufficient conditions for exact support recovery in the high-dimensional asymptotic regime.
This is a major theoretical contribution of the thesis, solving, and expanding on, open problems in the literature (see \cite{butucea2018variable, gao2018fundamental}).
The analysis of support recovery problem is intimately related to a \emph{concentration of maxima} phenomena in the analysis of extremes.
The latter concept is key to understanding the role played by dependence in the phase transition phenomena of high-dimensional testing problems.
Using this probabilistic concept, we establish minimax optimality results that hold for a very large class of dependence structures.

>>>>

A common theme in such applications is that the errors are correlated, and that the signal vectors are believed to be sparse: the number of non-zero components in $\mu$ is small compared to the number of test performed --- in the cybersecurity example, while monitoring is performed over a large number of servers, it is often believed that only very few servers will be experiencing problems at any time.
Under such sparsity assumptions, it is natural to ask if and when one can reliably {(1)} detect the signals, and {(2)} recover the support set $S$.
% Further, one would like to characterize the statistical procedures that achieve efficient detection and support recovery, whenever such goals become possible to achieve.
In this text, we explore both the \emph{detection} and the \emph{support recovery} problems.  More precisely, we are interested in the 
theoretical feasibility of both problems, and seek minimal conditions under which these problems can be consistently solved in large 
dimensions.

@@

Both the Gaussianity assumption and the independence assumption are relaxed in Chapter \ref{chap:exact-support-recovery}.
Established are the necessary and sufficient conditions for exact support recovery in the high-dimensional asymptotic regime.
This is a major theoretical contribution of our approach, which solves and expands on open problems in the recent literature 
(see \cite{butucea2018variable, gao2018fundamental}).
The analysis of support recovery problem is intimately related to a \emph{concentration of maxima} phenomena in the analysis of extremes.
The latter concept is key to understanding the role played by dependence in the phase transition phenomena of high-dimensional testing problems.
Using this probabilistic concept, we establish minimax optimality results that hold for a very large class of dependence structures.

### 2020-08-21 2.background/2.AGG.tex

We introduce a fairly general class of distributions known as \ac{AGG} in this section, and state some properties about the tails of the AGG distributions.

\begin{definition} \label{def:AGG}
A distribution $F$ is called asymptotic generalized Gaussian with parameter $\nu>0$ (denoted $\text{AGG}(\nu)$) if
\begin{enumerate}
    \item $F(x)\in(0,1)$ for all $x\in\R$, and \smallskip
    \item $\log{\overline{F}(x)} \sim -\frac{1}{\nu}x^\nu$ and $\log{F(-x)} \sim -\frac{1}{\nu}(-x)^\nu,$ \label{eq:AGG}
\end{enumerate}
where $\overline{F}(x) = 1 - F(x)$ is the survival function, and $a(x)\sim b(x)$ is taken to mean $\lim_{x\to\infty} a(x)/b(x) = 1$.
\end{definition}

>>>>

We introduce a fairly general class of distributions known as \ac{AGG} in this section, and state some properties about the tails of the 
\index{AGG}{AGG} distributions.

\begin{definition} \label{def:AGG}
A distribution $F$ is called \index{asymptotic generalized Gaussian}{asymptotic generalized Gaussian} 
with parameter $\nu>0$ (denoted $\text{AGG}(\nu)$) if
\begin{enumerate}
    \item $F(x)\in(0,1)$ for all $x\in\R$, and \smallskip
    \item $\log{\overline{F}(x)} \sim -\frac{1}{\nu}x^\nu$ and $\log{F(-x)} \sim -\frac{1}{\nu}(-x)^\nu,$ \label{eq:AGG}
\end{enumerate}
where $\overline{F}(x) = 1 - F(x)$ is the survival function, and $a(x)\sim b(x)$ is taken to mean $\lim_{x\to\infty} a(x)/b(x) = 1$.
\end{definition}

### 2020-08-21 2.background/2.Exercises.tex

 {\em Hint:} In part {\bf (a)}, argue that the events $\{ M_p > (1+\delta_p) u_p,\ \mbox{ infinitely often}\}$ and
  $\{\epsilon(p) > )(1+\delta_p)u_p, \ \mbox{infinitely often}\}$ are equal. Appeal to the Borel Zero-One Law.\\

>>>>

 {\em Hint:} In part {\bf (a)}, argue that the events $\{ M_p > (1+\delta_p) u_p,\ \mbox{ infinitely often}\}$ and
  $\{\epsilon(p) > (1+\delta_p)u_p, \ \mbox{infinitely often}\}$ are equal. Appeal to the Borel Zero-One Law.\\

### 2020-08-21 acronyms.tex

 \begin{description}[CABR]
  \item[AGG]{asymptotically generalized Gaussian}
  \item[BH]{Benjamini-Hochberg}
 \item[CDF]{cumulative distribution function}
 \item[FDR]{false discovery rate}
 \item[FNR]{false non-discovery rate}
 \item[FWER]{family-wise error rate}
 \item[FWNR]{family-wise non-discovery rate}
 \item[GWAS]{genome-wide association studies}
 \item[HC]{higher-criticism}
 \item[iid]{independent and identically distributed}
 \item[LR]{likelihood-ratio}
 \item[RS]{relatively stable/relative stability}
 \item[SNP]{single-nucleotide polymorphisms}
 \item[URS]{uniform relatively stable/uniform relative stability}
 \end{description}

>>>>

 \begin{description}[CABR]
  \item[\index{AGG}{AGG}]{asymptotically generalized Gaussian}
  \item[\index{BH}{BH}]{Benjamini-Hochberg}
 \item[\index{CDF}{CDF}]{cumulative distribution function}
 \item[\index{FDR}{FDR}]{false discovery rate}
 \item[\index{FNR}{FNR}]{false non-discovery rate}
 \item[\index{FWER}{FWER}]{family-wise error rate}
 \item[\index{FWNR}{FWNR}]{family-wise non-discovery rate}
 \item[\index{GWAS}{GWAS}]{genome-wide association studies}
 \item[\index{HC}{HC}]{higher-criticism}
 \item[\index{iid}{iid}]{independent and identically distributed}
 \item[\index{LR}{LR}]{likelihood-ratio}
 \item[\index{RS}{RS}]{relatively stable/relative stability}
 \item[\index{SNP}{SNP}]{single-nucleotide polymorphisms}
 \item[\index{URS}{URS}]{uniform relatively stable/uniform relative stability}
 \end{description}
